---
title: "Apple Health Notebook"
output: 
  html_document: 
    toc: yes
editor_options: 
  markdown: 
    wrap: 72
---

# Apple Health Notebook

This notebook assumes that you have created and extracted an export with
your Apple Health data in the current working directory. It is highly
inspired by [Taras
Kaduk](https://taraskaduk.com/posts/2019-03-23-apple-health/).

## Load required libraries and xml data

```{r}
library(XML)
library(tidyverse)
library(lubridate)
library(scales)
library(ggthemes)
setwd('./apple_health_export')
xml <- xmlParse('Export.xml')
```

## Create dataframes

```{r}
df_record   <- XML:::xmlAttrsToDataFrame(xml["//Record"])
df_activity <- XML:::xmlAttrsToDataFrame(xml["//ActivitySummary"])
# df_workout  <- XML:::xmlAttrsToDataFrame(xml["//Workout"])
# df_clinical <- XML:::xmlAttrsToDataFrame(xml["//ClinicalRecord"])
# df_location <- XML:::xmlAttrsToDataFrame(xml["//Location"])
```

## Cleanup data

-   **`Record`** is the main place where the data is stored. Weight,
    height, blood pressure, steps, nutrition data, heart rate - all
    stored here

-   **`ActivitySummary`** is your Apple Watch daily Activity stats:
    Move, Exercise, Stand data

-   **`Workout`** is your Apple Watch workout activity per workout
    logged

-   **`Location`** is your location logged during your Apple Watch
    workouts (useful for runs/hikes)

-   **`InstantaneousBeatsPerMinute`** is exactly that: instantaneous
    heart rate when measured by AppleWatch

-   **`ExportDate`** is useful to validate what data are you looking at.

### Record

The record dataframe has some interesting vectors:

-   `sourceName` and `device` - where the data was generated, e.g. the
    name "John's Apple Watch" as `sourceName` and "Apple Watch" as the
    `device`

-   `type` - the type of record, could be `BodyMass` for weight or
    `VO2Max` for respitory fitness

-   `value` - the actual value with more details in the `unit` factor

```{r}
df_record <- df_record %>%
mutate(device = gsub(".*(name:)|,.*", "",device),
       value = as.numeric(as.character(value)),
       endDate = ymd_hms(endDate,tz="Europe/Berlin"),
       date = date(endDate),
       year = year(endDate),
       month = month(endDate),
       day = day(endDate),
       yday = yday(endDate),
       wday = wday(endDate),
       hour = hour(endDate),
       minute = minute(endDate),
       type = str_remove(type, "HKQuantityTypeIdentifier")
       )

df_record %>% 
  select(type) %>% 
  distinct()
```

### Activity

The activity data frame has several interesting factors:

-   `dateComponents` - the actual date for the activity stats

-   `activeEnergyBurned` - the actual number of calories burned on this
    day (`activeEnergyBurnedUnit` typically is set to kcal, hence it is
    less interesting to me)

-   `activeEnergyBurnedGoal` - the calories goal for this day

-   `appleExcerciseTime` and `appleExcerciseTimeGoal` - the actual and
    goal for activity minutes on a day

-   `appleStandHours` and `appleStandHoursGoal` - actual and goal for
    hours containing at least 60 secs of standing

```{r}
df_activity <- df_activity %>%
mutate(date = as.date(dateComponents),
       year = year(date),
       month = month(date),
       day = day(date),
       yday = yday(date),
       wday = wday(date)
       )

```

## Visualize data

### Plotting weight over time

```{r}
df_record %>%
  arrange(endDate) %>% 
  filter(type == 'BodyMass') %>% 
  filter(date > '2023-01-01' ) %>% 
  
  ggplot(aes(x= date, y = value)) +
    geom_point(alpha = 0.3) +
    geom_smooth(span = 0.2, col = "grey30", se = FALSE) +
    labs(title = "Apple Health Weight Chart",
         caption = "powered by ggplot") +
    xlab("Date") + ylab("Weight")
```

## Plotting activity/step count

```{r}
df_record %>%
  filter(type == 'StepCount') %>% 
  group_by(date,wday,hour) %>% 
  summarize(steps=sum(value)) %>% 
  group_by(hour,wday) %>% 
  summarize(steps=sum(steps)) %>% 
  arrange(desc(steps)) %>%

  ggplot(aes(x=hour, y=wday,  fill=steps)) + 
    geom_tile(col = 'grey40') + 
    scale_fill_continuous(labels = scales::comma, low = 'grey95', high = '#008FD5') +
    theme(panel.grid.major = element_blank()) +
    scale_x_continuous(
      breaks = c(0, 6, 12, 18),
      label = c("Midnight", "6 AM", "Midday", "6 PM")
    ) +
    scale_y_reverse(
      breaks = c(1,2,3,4,5,6,7),
      label = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
    ) +
    labs(title = "Weekly Step Count Heatmap",
         caption = 'powered by ggplot') +
    guides(fill=FALSE)+
    coord_equal()
```

### Plot activity

```{r}
df_activity %>%
  arrange(date) %>% 
  
  ggplot(aes(x= date, y = activeEnergyBurned)) +
    geom_point(alpha = 0.3) +
    geom_smooth(span = 0.2, col = "grey30", se = FALSE) +
    labs(title = "Apple Health Energy Burned Chart",
         caption = "powered by ggplot") +
    xlab("Date") + ylab("Weight")
```

## Activity Data

```{}
```
