---
title: "Apple Health Notebook"
output: 
  html_document: 
    toc: yes
editor_options: 
  markdown: 
    wrap: 72
---

# Apple Health Notebook

This notebook assumes that you have created and extracted an export with
your Apple Health data in the current working directory. It is highly
inspired by [Taras
Kaduk](https://taraskaduk.com/posts/2019-03-23-apple-health/).

## Load required libraries and xml data

```{r}
library(XML)
library(tidyverse)
library(lubridate)
library(scales)
library(ggthemes)
setwd('./apple_health_export')
xml <- xmlParse('Export.xml')
```

## Create dataframes

```{r}
df_record   <- XML:::xmlAttrsToDataFrame(xml["//Record"])
df_activity <- XML:::xmlAttrsToDataFrame(xml["//ActivitySummary"])
# df_workout  <- XML:::xmlAttrsToDataFrame(xml["//Workout"])
# df_clinical <- XML:::xmlAttrsToDataFrame(xml["//ClinicalRecord"])
# df_location <- XML:::xmlAttrsToDataFrame(xml["//Location"])
```

## Cleanup data

-   **Record** is the main place where the data is stored. Weight,
    height, blood pressure, steps, nutrition data, heart rate - all
    stored here

-   **ActivitySummary** is your Apple Watch daily Activity stats: Move,
    Exercise, Stand data

-   **Workout** is your Apple Watch workout activity per workout logged

-   **Location** is your location logged during your Apple Watch
    workouts (useful for runs/hikes)

-   **InstantaneousBeatsPerMinute** is exactly that: instantaneous heart
    rate when measured by AppleWatch

-   **ExportDate** is useful to validate what data are you looking at.

### Record

```{r}
df_record <- df_record %>%
mutate(device = gsub(".*(name:)|,.*", "",device),
       value = as.numeric(as.character(value)),
       endDate = ymd_hms(endDate,tz="Europe/Berlin"),
       date = date(endDate),
       year = year(endDate),
       month = month(endDate),
       day = day(endDate),
       yday = yday(endDate),
       wday = wday(endDate),
       hour = hour(endDate),
       minute = minute(endDate),
       type = str_remove(type, "HKQuantityTypeIdentifier")
       )

df_record %>% 
  select(type) %>% 
  distinct()
```

## Visualize data

### Plotting weight over time

```{r}
df_record %>%
  arrange(endDate) %>% 
  filter(type == 'BodyMass') %>% 
  filter(date > '2023-01-01' ) %>% 
  
  ggplot(aes(x= date, y = value)) +
    geom_point(alpha = 0.3) +
    geom_smooth(span = 0.2, col = "grey30", se = FALSE) +
    labs(title = "Apple Health Weight Chart",
         caption = "powered by ggplot") +
    xlab("Date") + ylab("Weight")
```

## Plotting activity/step count

```{r}
df_record %>%
  filter(type == 'StepCount') %>% 
  group_by(date,wday,hour) %>% 
  summarize(steps=sum(value)) %>% 
  group_by(hour,wday) %>% 
  summarize(steps=sum(steps)) %>% 
  arrange(desc(steps)) %>%

  ggplot(aes(x=hour, y=wday,  fill=steps)) + 
    geom_tile(col = 'grey40') + 
    scale_fill_continuous(labels = scales::comma, low = 'grey95', high = '#008FD5') +
    theme(panel.grid.major = element_blank()) +
    scale_x_continuous(
      breaks = c(0, 6, 12, 18),
      label = c("Midnight", "6 AM", "Midday", "6 PM")
    ) +
    scale_y_reverse(
      breaks = c(1,2,3,4,5,6,7),
      label = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
    ) +
    labs(title = "Weekly Step Count Heatmap",
         caption = 'powered by ggplot') +
    guides(fill=FALSE)+
    coord_equal()
```

### Combine weight and steps

```{r}
#TODO: Write pretty ggplot code for bar representing steps and line showing weight
```

## Activity Data

```{r}
df_activity_clean <- df_activity %>% 
  select(-activeEnergyBurnedUnit) %>% 
  mutate_all(as.character) %>% 
  mutate(date = as.Date(dateComponents)) %>% 
  filter(date >= "2018-11-13") %>% 
  select(-dateComponents) %>% 
  mutate_if(is.character, as.numeric) %>% 
  rename(move = activeEnergyBurned,
         exercise = appleExerciseTime,
         stand = appleStandHours,
         move_goal = activeEnergyBurnedGoal,
         exercise_goal = appleExerciseTimeGoal,
         stand_goal = appleStandHoursGoal) %>% 
#Now, create 2 new metrics: percent of goal and a "Yes/No" flag.
  mutate(move_pct = move/move_goal,
         exercise_pct = exercise/exercise_goal,
         stand_pct = stand/stand_goal,
         move_bool = if_else(move_pct < 1, FALSE, TRUE),
         exercise_bool = if_else(exercise_pct < 1, FALSE, TRUE),
         stand_bool = if_else(stand_pct < 1, FALSE, TRUE))

```
