---
title: "R Timeseries Data"
output: html_notebook
---

# This notebook deals with data over a period of time.

## Sample data

Our sample data is present in `data/data.csv`. It contains a simple timestamp and two values.

## Load libraries

```{r}
prereqs<-c("tidyverse", "lubridate")
lapply(prereqs, require, character.only = TRUE)
```

## Read data

We're reading in the csv file and assign better column names while we're at it.

```{r}
df <- read.csv(file = 'data/data.csv')

# assign meaningful column names
colnames(df) <- c("timestamp",
  "value1",
  "value2")

head(df)
```

## Add usable timestamps

[lubridate](https://lubridate.tidyverse.org) gives several nice tools to manipulate dates and extract valuable information, like the weekday or a proper timestamp.

```{r}
df <- mutate(df,
  ymdhms = ymd_hms(timestamp),
  date = date(timestamp),
  day_of_week = wday(date),
  weekday_long = strftime(date,'%A'),
  weekday_short = strftime(date,'%a'),
  month = month(date),
  week = week(date)
)
  
head(df)
```

## A first plot

Using ggplot. Check out [this cheat sheet](https://github.com/rstudio/cheatsheets/blob/main/data-visualization-2.1.pdf).

```{r}
p <- ggplot(data = df, aes(x = date, y = value1))+
  geom_line(color = "#00AFBB", linewidth = 2)

p + stat_smooth(
  color = "#FC4E07", fill = "#FC4E07",
  method = "loess"
  )
```

Another variant

```{r}
library(scales)
ggplot(df, aes(x = date, y = value1)) + 
  geom_bar(stat = "identity") +
  theme_bw() +
  labs(x = "Date", y = "value") +
  scale_x_date(
    labels = date_format("%b\n%Y"), 
    date_breaks = "6 months")
```

## Grouping by Months

```{r}
df_ausgaben <- read.csv(file = 'data/ausgaben.csv', sep=';')

df_ausgaben <- mutate(df_ausgaben,
  Datum = as.Date(Datum, format = "%d.%m.%Y"),
  Monat = month(Datum),
  Jahr = year(Datum),
  Betrag =as.numeric(gsub(",", ".", gsub("\\.", "", Betrag)))
)
  
head(df_ausgaben)

```

Now for the monthly expenses

```{r}
df_ausgaben_month <- df_ausgaben %>%
  group_by(Monat) %>%
  summarise(Ausgaben = sum(Betrag))

df_ausgaben_month
```
